### 2014-02-25 doc/proposal/proposal.Rmd
Proposal draft

intRo
========================================================

### Motivation 
To create a web-based tool to be used in introductory statistics class in lieu of JMP. Ultimately, our goal is to create an interest in coding while learning basic statistics, or at a minimum, help students to learn a bit more about working with data.

As a web-based application, this tool will be immediately more familiar to students than a desktop application. The need for dealing with software licenses, installation configuration, and supported platforms will be gone. This will allow students to spend more time working with the data and learning statistics than having to struggle to get the software running.

The application will combine several different modules which perform specific data analysis tasks.

### Modules
* Data Import (i.e., displaying data, subsets...)
* Descriptive Summaries (i.e., mean, median, variance, IQR...)
* Graphical Summaries (i.e., bar chart, pareto chart, scatterplot, histogram...)
* Statistical Tests (i.e., t test, z test...)
* Simulation (i.e., central limit theorem...)
* Linear Regression (i.e., fitting, plotting, least squares, prediction...)

Students will be able to view the R code that each module generates. When changing configuration options in each module, the students will be able to observe the specific changes occurring in the R code. Students will be able to import their own dataset, or work with datasets that are provided in the application.

### Class Deliverables
We aim to have a prototype of this application with Data Import, Descriptive Summaries, and Graphical Summaries completed for the class. We will include accompanying documentation for using intRo.