### 2014-12-16 hmwkch4.tex
first 6 homeworks plus datasets

\documentclass{article}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{epsfig, amsmath, amssymb}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}

\hspace{2.5in} Name: (Print) \underline{\mbox{\hspace{3in}}}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 4}\\[3mm]
\end{center}

\begin{enumerate}

\item One hundred middle-aged men and one hundred young adult men were rated on physical flexibility.  The scale went from 1, least flexible to 10, most flexible.  The table below shows a summary of these ratings.  For example, only 1 Young Adult Man had a flexibility rating of 1 while 2 Middle-Aged Men received this rating.

\begin{center}
\begin{tabular}{l|cccccccccc} \hline
Flexibility Rating & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\ \hline
Young Adult Men & 1 & 3 & 6 & 8 & 11 & 14 & 16 & 19 & 14 & 8\\
Middle-Aged Men & 2 & 6 & 10 & 14 & 17 & 18 & 14 & 11 & 6 & 2\\ \hline
\end{tabular}
\end{center}

Calculate the 5-number summary for the flexibility ratings for Young Adult Men and for Middle-Aged Men.  Make sure to clearly label the values and the group they belong to.

\newpage

\item A study examining the health risks of smoking measured the cholesterol levels of people who had smoked for at least 25 years and people of similar ages who had smoked for no more than 5 years and then stopped.  The cholesterol levels and smoking status of all 76 people can be found in the JMP data file {\bf Cholesterol.JMP}  in Blackboard.

\begin{enumerate}
\item Describe the Who for this data set.\\[0.5in]
\item Describe the What for this data set.\\[0.5in]
\item Use JMP to obtain the distribution of cholesterol levels for both groups (Smokers and Non-Smokers).  Print out this output and turn it in with your assignment.\\[1em]
\item Use the means and medians of the two distributions to discuss similarities and/or differences in the centers of the two distributions.\\[1in]
\item Use the ranges, IQRs, and standard deviations of the two distributions to discuss similarities and/or differences in the variability of the two distributions.\\[1in]
\item Doctors want patients to maintain a cholesterol level less than 200 points.  What percent of people in the Smokers group reached this goal?  What percent of people in the Ex-Smokers group reached this goal?  Why would it be important to report this information in the analysis of these data?\\[0.5em]

Percent in Smokers group less than 200 = \underline{\mbox{\hspace{1.5in}}}\\[1.5em]
Percent in Ex-Smokers group less than 200 = \underline{\mbox{\hspace{1.5in}}}\\[1em]

Why important to report?:
\end{enumerate}

\end{enumerate}

\end{document}