### 2014-12-16 hmwkch5.tex

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}
\hspace{2in} Name: (Print) \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 5}\\[3mm]
\end{center}

\begin{enumerate}

\item The GRE is a test to measure undergraduate achievement and is used as an admissions criterion to graduate school. The GRE is designed in such a way that scores on the test can be modeled with a normal model. From July 2006 to June 2007 the mean GRE Verbal score was 457 with a standard deviation of 121. Use the z-table to answer the following questions.

\begin{enumerate}
\item What proportion of all test takers will have a GRE Verbal score above 600 points?\\[1in]
\item What proportion of all test takers will have a GRE Verbal score below 375 points?\\[1in]
\item What proportion of all test takers will have a GRE Verbal score between 500 and 700 points?\\[1in]
\item What GRE Verbal score will 40\% of all test takers score below?\\[1in]
\item What GRE verbal score will 10\% of all test takers score above?
\end{enumerate}

\newpage

\item In a study to determine the nature of the relationship between heart rates and body temperature, body temperatures were collected from 130 individuals. The body temperatures, in degrees Fahrenheit, can be found in the JMP data file {\bf samplebodytemps.JMP} in Blackboard.

Use JMP to determine if the distribution of body temperatures can be modeled with a normal model. Make sure to turn in the output from JMP with your assignment. In your answer, make sure to use information from the JMP output that supports your conclusions.

\end{enumerate}

\end{document}