### 2015-01-06 departmental-seminar/seminar-slides-markdown.Rmd
working with rmarkdown instead

---
title: "`intRo`"
subtitle: "Statistical Analysis Software for Teaching"
author: "Andee Kaplan & Eric Hare"
date: "January 12, 2015"
output: ioslides_presentation
font-family: Helvetica
ioslides_presentation:
    logo: images/icon.png
    widescreen: true
---

# Introduction

## Do we really need another statistical software package?
Short answer: **yes**

 - `R` is great, but requires students to have some knowledge/interest in programming
 - JMP is powerful, but confusing to students
   - Licenses
   - Inconsistent UI
 - Students fight with software rather than focusing on the goal: statistics

## What is `intRo`?
 - A modern **web-based** application for performing basic data analysis and statistical routines
 - Built using `R` and `Shiny`
 - Powerful and extensible modular structure
 - Simple enough for the novice statistician
 - Meant to assist in the learning of statistics rather than a stand-alone deliverer of statistics education

## Easy
- Focused on aspects of the user interface (UI) and output that make it easy to pick up without training
- Minimal necessary functionality for an introductory statistics course
- Organized around specific tasks a student may perform in the process of a data analysis

<img src=images/user_experience.png height="250"></img>

## Exciting
 - Fun, easy to use (available on the web)
 - Interactive plots using `ggvis`

**Ulterior motive**: get students excited about programming
 - By navigating about the user interface of `intRo`, students are creating a fully-executable `R` script that they can download and run locally
 - Viewing their script change real-time within the application

## Extensible
 - User interaction with `intRo` is split into bitesize chunks that we call *modules*
 - Each module is a self contained set of `R` code that is dynamically added to the application at run time
 - `intRo` can be easily extended by the addition of modules within the frame-work underlying the application
 - Allows `intRo` to be tailored to the needs of a particular course

# First look | http://intro-stats.com

# Classroom

## `intRo` in the wild
 - `intRo` is currently being used in STAT 201
  - Homework assignments and labs will use `intRo` to emphasize class concepts
 - `intRo` is a supplement, not the focus of the class

## Homework 1
 - *Data*: OECD PISA Results in Focus report to assess the workforce readiness of 15-year old students
 - *Goal*: Explore graphically and numerically how the students perceive academics and math
 - `intRo` *Tasks*:
  1. Make a summary of the Enjoy.Maths variable. What proportion of these students agree or strongly agree with the statement "I enjoy mathematics"?
  1. Make a barchart of the variable Enjoy.Maths.
  1. Give the conditional distribution of "Strongly agree'' given the Country.
  2. Obtain a contingency table and mosaic plot of the relationship between Country and Enjoy.Maths

# Our turn

# Design decisions

## Consistent UI
Elements of `intRo`: 1) top navigation, 2) side navigation, 3) options
panel, 4) results panel, and 5) code panel.

<img src="images/ui_annotate.png" height="400"></img>

## Consistent UI (Cont'd)
 - Large, easy to click icons in the page header to help students find
exactly what they need easily
 - Each module maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code
 - Documentation website that is consistent with
the interface of the application; covers all default modules and makes learning to use `intRo` even more painless

## Modularity
<img src="images/app_creation_modules.png" height="500"></img>

## Modularity (Cont'd)
The interface is created with the following statement

```r
## Source ui
module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)

## mylist is a list containing the different ui module code
## Create the UI
shinyUI(
    navbarPage("intRo", id="top-nav",  theme = "bootstrap.min.css",
               tabPanel(title="", icon=icon("home"),
                        fluidRow(
                        ### <b>
                            do.call(navlistPanel, c(list(id = "side-nav", widths = c(2, 10)), mylist))
                        ### </b>
                        )
               ), ...
    ))
```

## Modularity (Cont'd)
The server functions are dynamically generated using a similar method

```r
shinyServer(function(input, output, session) {
  ## Module info
  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")

  ## Modules
  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
  modules_tosource <- file.path("modules", 
                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))

  ## Source the modules
  for (mod in modules_tosource) {
      source(mod, local = TRUE)
  }
})
```

# Challenges

## Tradeoffs
There were a number of challenges we encountered in striving to strike the balance of functionality and ease-of-use:

 - Storing the user's results
 - Presenting only relevant options to the user
 - Producing clean and executable R code

## Storing results
 - Users must be able to print results, but only the results they want
 - Reactivity makes this somewhat challenging

**Solution**: Store Buttons

<img src="images/store.png" height="300"></img>

## Storing results (Cont'd)
 - Each module contains, at the bottom, a Store button
 - When Store is pressed, code is stored in code panel
     - Additional challenge: All options selected must also be stored!
 - When user goes to print results, only this particular code is executed

## Presenting choices
 - `intRo` is flexible, but must also be simple for the introductory student
 - Despite increasing the complexity of the code, we decided to only show options appropriate for the given selections.

**Example**:

<img src="images/graphical.png" height="300">

## Presenting choices (Cont'd)
<img src="images/graphical_table.png" height="275">

## Producing clean code
Code generated must be:

 - Executable on the server
 - Executable on the user's machine
 - Clean

## Producing clean code (Cont'd)
But how do we handle uploaded data?

```{r, eval=FALSE}
file.choose <- function () {
    return(input$data_own[,"datapath"])
}
```

Now we can provide to the user *and* the server:

```{r, eval=FALSE}
intro.data <- read.csv(file.choose())
```

# Future work

## What's next?
 - Module creation package
 - Server load
 - Use in other classes

## Module creation package
*Modularity* is a key feature of `intRo`, but module creation is currently:
 - Undocumented
 - Entirely manual
 - Unnecessarily lengthy

**Idea**: R package (with associated Shiny app?) to automate creation of `intRo` modules

## Server load
Limited testing has been done to assess how `intRo` handles heavy user load
 - Dozens of students at accessing app at once
 - Simultaneous larger computations like printing results

More server resources may need to be devoted to handle this, particularly if `intRo` is more widely adopted

## More classes
 - We would like to see `intRo` used in more classes
 - Module creation package will allow it to suit more curricula
 - We welcome collaborators interested in extending `intRo` to submit pull requests on GitHub (http://www.github.com/gammarama/intRo)
     - Additional modules
     - Improvements and bug fixes

## Special thanks
Dr. Cook, for guiding many of the features and interface decisions, and helping test the functionality in preparation for STAT 201

# Any questions? | Thank you!

Contact:<br/>
ajkaplan@iastate.edu<br/>
erichare@iastate.edu