### 2015-02-16 lessonplans/hmwkch3.tex
homeworks to be done using intRo

\documentclass{article}[11pt]
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}

\hspace{2in} Name: (Print) \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\Large STATISTICS 201 - Written Homework Chapter 3}\\[3mm]
\end{center}

\begin{enumerate}

\item Michael Jordan is highly renowned basketball player who played for 15 seasons in the NBA.  The following table contains the average points per game scored by Jordan in each of his 15 seasons.

\begin{center}
\begin{tabular}{|l|c||l|c||l|c|} \hline
Year & Avg Pts/Game & Year & Avg Pts/Game & Year & Avg Pts/Game 
\\ \hline
1984 & 28 & 1989 & 34 & 1995 & 30 \\
1985 & 23 & 1990 & 32 & 1996 & 30 \\
1986 & 37 & 1991 & 30 & 1997 & 29 \\
1987 & 35 & 1992 & 33 & 2001 & 23 \\
1988 & 33 & 1994 & 27 & 2002 & 20 \\ \hline
\end{tabular}
\end{center}

\begin{enumerate}
\item Make a stem and leaf display of the average number of points per game that Michael Jordan scored per season.  Do this by hand, use a split stem, and ensure that the leaves are sorted numerically. Describe the shape of this distribution. \\[2.5in]

\item Make a line plot of average number of points per game by year. Describe the pattern of points relative to time. \\[2.5in]

\end{enumerate}

\newpage
\item For this part of the assignment you will be using data collected on IA high school quarterbacks for the 2012 season.

\begin{enumerate}
\item Describe the Who for this data.\\[0.35in]
\item Use intRo to obtain the a histogram of yards thrown, with a suitable bin width.  Describe the shape of the distribution of this variable. Print out the output and turn it in with this assignment.
\item Give the five number summary for this variable.\\[0.2in]
\item Calculate the range and IQR for this variable.\\[0.1in]
\item What percentage of the quarterbacks  threw less than 366 yards? \\[0.1in]
\item What percentage of the quarterbacks  threw more than 3396 yards?  \\[0.1in]
\item Give the mean and standard deviation for this variable.\\[0.1in]
\item Explain why there is a difference of approximately 70 yards in the mean and median for this variable.\\[1in]
\item Which numerical summaries are the most appropriate for this distribution - the 5-number summary or the mean and standard deviation?  Circle the appropriate answer below.  Then explain your answer.

\hspace{1in} 5-number summary \hspace{1in} mean and standard deviation \\[0.5em]

Explain your answer:\\[0.5in]
\item Explore transformations of this variable to make it more symmetric. Circle which is best of the choices below:

\hspace{1in} 0.5 \hspace{1in} 0.25 \hspace{1in} 0

Print a histogram of your transformed data. Compute what the transformed value of 3000 yards would be.

\end{enumerate}

\end{enumerate}

\end{document}