### 2015-02-16 lessonplans/hmwkch6-7.tex

\documentclass{article}
\oddsidemargin 0in \evensidemargin 0in \topmargin 0in
\columnsep 10pt \columnseprule 0pt 
\marginparwidth 90pt \marginparsep 11pt \marginparpush 5pt 
\headheight 0pt \headsep 0pt 
%\footheight 0pt \footskip 0pt 
\textheight 9.0in \textwidth 6.5in

\begin{document}
\hspace{2in} Name: \underline{\mbox{\hspace{3.5in}}}\\

\begin{center}
{\large STATISTICS 201 - Written Homework Chapter 6}\\[3mm]
\end{center}

\begin{enumerate}
\item For the Somerset neighborhood 2009 Ames sales data tabulate the correlations between the SalePrice, LotArea, LivingArea and GarageArea. (See the textbook for an example of how to do this. Be sure to round numbers appropriately.)

%\vspace{2.5in}

\begin{enumerate}
\item Which two variables have the strongest correlation? What is it?

%\vspace{1in}

\item Make scatterplots of each of the pairs of variables (don't hand these in). Write a paragraph briefly describing the associations, and discussing whether correlation is  or isn't an appropriate summary.
\end{enumerate}

\item Fit a regression model for SalePrice against LivingArea.

\begin{enumerate}
\item Write out the model. 
\item Print the scatterplot and sketch the linear model on this plot. 
\item What proportion of variation in house price does the living area explain?
\item
Predict the average sales price for a house having 1500 sq ft of living space.
\item There is a (small) problem with the fit. Diagnose this using the residual plots, and write a paragraph on how you would correct it, and re-fit the model. 
\end{enumerate}
\end{enumerate}

\end{document}