### 2016-06-06 jcgs-paper-2016/cover_letter.Rmd
cover letter template

---
author: 
 - Eric Hare
 - PhD Student 
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: To whom it may concern,
closing: Best regards,
address: 
 - Editor
 - Journal of Computational and Graphical Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

Thank you in advance for reviewing our manuscript "Introductory statistics with `intRo`", which I have submitted for consideration of publication in the Journal of Computational and Graphical Statistics.

This paper details an interactive web-based statistical software application intended for use in an introductory statistics classroom. We introduce the tool as well as an accompanying R package to ease the process of deploying intRo within a curriculum. Additionally, the paper describes the process for extending the functionality in intRo in order to personalize the software.

Thank you for your consideration of our submission.