### 2016-10-20 jcgs-paper-2016/cover_letter-1strev.Rmd
New cover letter

---
author: 
 - Eric Hare
 - PhD Student 
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: Dear Dr. Cook,
closing: Best regards,
address: 
 - Dianne Cook
 - Editor
 - Journal of Computational and Graphical Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

Re: Manuscript reference No. JCGS-16-045

Thank you for the reviewers' comments and the opportunity to resubmit our manuscript, "Introductory statistics with intRo". We feel the paper is stronger as a result of incorporating the feedback.

We look forward to hearing your thoughts on our changes, and we appreciate your consideration of our resubmission.