### 2016-10-20 jcgs-paper-2016/cover_letter.Rmd

---
author: 
 - Eric Hare
 - PhD Student 
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: Dear Dr. Cook,
closing: Best regards,
address: 
 - Dianne Cook
 - Editor
 - Journal of Computational and Graphical Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

Re: Manuscript reference No. JCGS-16-045

Attached you will find the newest revision of our paper, written in response to the feedback we received from the reviewers, the associate editor, and yourself.

We have attempted to restructure the paper to focus on general techniques for designing a modular software system, using intRo itself as a case study. In particular, we have re-written the title, abstract, and a substantial portion of the background to highlight this new direction.

In addition to the refocusing, we have also extended our review of other software systems, and emphasized the need that intRo addresses. Table 1 provides a somewhat subjective but thorough comparison of many "competing" software systems in terms of a number of qualities, including whether or not they are web-based, the level of extensibility, and the amount of functionality. We have also done our best to address the other minor comments that were raised in the second round of reviews.

We hope that this latest revision addresses the majority of the issues with the previous, and we welcome any further critiques or comments.