### 2016-12-09 jcgs-paper-2016/response.Rmd
New response and small edits

---
title: "Response to third review"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---

# Associate Editor

1. two brief comments from one reviewer should both be addressed

    **We have addressed these two comments**

2. from other reviewer, address the minor fixes

    **As noted in the comments in this document, we have addressed what we felt were the more minor fixes**

3. remove the comments about Radiant being "inundated with more features than necessary"

    **This comment has been removed**

4. a sentence or two more on comparison with Radiant, e.g., mention Radiant's less focus on *extensibility* (if that is true)

    **We have extended the comparison with Radiant by noting Radiant's more extensive functionality and the differences in the goals of each**

# Reviewer 2

5. Page 3 re iNZight Lite. The model fitting module in iNZight writes R reproducible code (and has since its inception) and it all will eventually so tone back this criticism a little. Given the sentence that comes before, I think this sentence could be just deleted without weakening the motivation for this paper and project.

    **We have clarified the differences in reproducibility as requested**

6. Figure 3: I take it that the contents of helper.R, libraries.R, ... get sucked in to populate/form server.R. This could be made clearer (or this misconception countered).

    **We have updated the figure caption to better explain how these files populate both the server and the ui definitions.**

# Reviewer 3

7. p1. Not sure when this package was originally written, but R Core Team 2014 probably needs to be updated to 2016.

    **We have updated the citation for R**

8. I don't follow the logic here. Perhaps give an example of which platform does not support them.

    **We have provided an example of this in the paper (Microsoft Excel and JMP are not supported on Linux)**

9. p2. "RStudio and Inc. 2014": the .bib entry should be author = {{RStudio, Inc.}}

    **Fixed**

10. perhaps change to "even if the instructors use R in their own work"?

    **Fixed**

11. p9. wilcoxon -> Wilcoxon (also on following pages)

    **We have corrected this throughout the paper**

12. p10. I think pseudocode suffices in Figure 3; no need to put actual code in it. In server.R, typically the arguments are input, output, session, and I'm not sure if session, input, output is the same.

    **We have updated the order of the arguments in figure 3.**

13. p15. Okay, interpolate() was explained in detail here. Then p12 should have referenced this section.

    **We have referenced this section on page 12**