### 2016-12-09 jcgs-paper-2016/cover_letter.Rmd
Final submission

---
author: 
 - Eric Hare
 - PhD Student 
 - Department of Statistics 
 - Iowa State University
date: "`r format(Sys.time(), '%B %d, %Y')`"
opening: Dear Dr. Cook,
closing: Best regards,
address: 
 - Dianne Cook
 - Editor
 - Journal of Computational and Graphical Statistics
return-address:
 - Eric Hare
 - Department of Statistics
 - Iowa State University
 - Ames, IA 50011
output:
  pdf_document:
    template: template-letter.tex
geometry: margin=1in
---

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

Re: Manuscript reference No. JCGS-16-045

Attached you will find the newest revision of our paper. Contained is:

- **paper.pdf**, the non-blinded version of our paper
- **paper_source.zip**, the source files for our paper, which include:
    - **paper.tex**, the tex source
    - **images**, a directory of figures
    - **refs.bib**, the references
    - **paper.pdf**, a copy of the non-blinded version of our paper

We also have included **forms.zip**, which has the two requested forms signed, as well as **response.pdf**, which has our final reviewer response.

Let us know if there are any further concerns or requests.