### 2019-07-28 jsm-2019/index.Rmd
started slides for jsm

---
title: "Designing Modular Software:"
subtitle: "A Case Study in Introductory Statistics"
author: "Eric Hare and **Andee Kaplan**"
institute: "Omni Analytics and Colorado State University"
date: "2019/07/29"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center, middle

# Background

### Some things to think about when writing software

---

# What is modularity?

### The degree to which a system's components may be separated and recombined.

#### Advantages

- Functionality can be dynamically loaded and unloaded depending on the **particular use case**

- Open source modularity allows **extensions** contributed by many people

- A **flexible system** that can be tailored to keep software easy to use

---

# What else is important?

### Reproducibility

Encourage **best practices** with regards to disclosure of analysis methods in research

### Extensibility

**Expands** into areas that weren't anticipated early in development

- e.g. packages in `R`, macros in Excel, add-ins in JMP

### Ease of use

Help users **avoid frustration**

- e.g., web technologies eliminate local installation, easy to learn GUI

---
class: center, middle

# Case study: `intRo`

### Statistical Analysis Software for Teaching

---

# Full disclosure

.pull-left[
![](images/attn.png)
<p style="font-size: 0.9rem;font-style: italic;"><a href="https://svgsilh.com/e91e63/image/98739.html">"warning"</a> is licensed under <a href="https://creativecommons.org/licenses/cc0/1.0/?ref=ccsearch&atype=html" style="margin-right: 5px;">CC0 1.0</a><a href="https://creativecommons.org/licenses/cc0/1.0/?ref=ccsearch&atype=html" target="_blank" rel="noopener noreferrer" style="display: inline-block;white-space: none;opacity: .7;margin-top: 2px;margin-left: 3px;height: 22px !important;"><img style="height: inherit;margin-right: 3px;display: inline-block;" src="https://search.creativecommons.org/static/img/cc_icon.svg" /><img style="height: inherit;margin-right: 3px;display: inline-block;" src="https://search.creativecommons.org/static/img/cc-cc0_icon.svg" /></a></p>
]

.pull-right[
We have an **ulterior motive** to get students excited about programming!

- By navigating about the user interface of `intRo`, students are creating a fully-executable `R` script that they can download and run locally

- Viewing their script change real-time within the application

- This is a sneaky way to **spark curiosity** of students

]

---
class: center, inverse, middle

# Live demo

### http://intro-stats.com

---

# Design decisions

---

# Challenges (that have now been solved)

---

# Get started

---

class: center, middle

# Thanks!

http://www.intro-stats.com

.footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).]